I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-15 Paper/Paper_noProof.tex
overall pass

		\end{tikzpicture}
		\caption{A DNN example from \cite{kneuron}, where every node is separated into 2, $x$ after the weighted sum and $\hat{x}$ after the ReLU activation function.}
		\label{fig1}
	\end{figure}

@@

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $weight(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (a contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of the exact $-x_2 \leq 1.5$), and obtain a bound $weight(x_5) \leq 5.5$ instead of $4$.

@@

This result has a first interesting application: explain why some DNNs are intrinsically easier to verify than others. This is the case for DNNs trained adversarially to be robust, vs same architecture, but trained in a "natural" way \cite{deeppoly,prima,crown}. Indeed, compensating pairs is a purely structural property on the DNN learnt, rather than more semantical notions, such as the number of unstable ReLUs, which depends heavily upon the image considered and the algorithm used. We compare compensations with source and target separated by 1 ReLU layer, e.g. source in layer 0 and target in layer 2, for two Networks with 5 hidden layers of 100 nodes each from the ERAN GitHub repository, one trained naturally ("$6\times100$"), 
and one trained using DiffAI ("$5\times100$"). 
We report in Table \ref{tab:compensation} the average compensation strength, i.e. 
the average of $max_{\pi,\pi'} min(|weight(\pi')|,|weight(\pi')|)$
over all source/target states in given layers. The average compensations strenght are much lower in the Robust DNN than in the Natural DNN, and indeed the Robust DNN is much simpler to verify than the Natural DNN.

@@

Notice that It is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus didnot experiment.

@@

		$5 \times100$\  &   $16\%$ & $51\%$ & $\mathbf{69.9\%}$ & no & point & $65.3\%$ & $68.4\%$ &  $84.2 \%$ \\ 
		$\epsilon = 0.026$ & 5s & 159s & 102s& & & 117s & 142s &  \\
		\hline

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accuate than using compensation: we are within $1.5\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).

>>>>

		\end{tikzpicture}
		\caption{A DNN example from \cite{kpoly}, where every node is separated into 2, $x$ after the weighted sum and $\hat{x}$ after the ReLU activation function. The pair $(x_2 x_3 x_5,x_2 x_4 x_5)$ in green and red is compensating (weights of paths are $1,-2$).}
		\label{fig1}
	\end{figure}

@@

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $weight(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of exact $-x_2 \leq 1.5$), obtaining a bound $weight(x_5) \leq 5.5$ instead of $4$.

@@

This result has a first interesting application: explain why some DNNs are intrinsically easier to verify than others. This is the case for DNNs trained adversarially to be robust, vs same architecture, but trained in a "natural" way \cite{deeppoly,prima,crown}. Indeed, compensating pairs is a purely structural property on the DNN learnt, rather than more semantical notions, such as the number of unstable ReLUs, which depends heavily upon the image considered and the algorithm used. We compare compensations with source and target separated by 1 ReLU layer, e.g. source in layer 0 and target in layer 2, for two Networks with 5 hidden layers of 100 nodes each from the ERAN GitHub repository, one trained naturally ("$6\times100$"), 
and one trained using DiffAI ("$5\times100$"). 
We report in Table \ref{tab:compensation} the average compensation strength, i.e. 
the average of $max_{\pi,\pi'} min(|weight(\pi')|,|weight(\pi')|)$
over all source/target states in given layers. The average compensations strenght are much lower in the Robust DNN than in the Natural DNN, and indeed the Robust DNN is much simpler to verify than the Natural DNN (e.g. DeepPoly is more accurate and verifies more images).

@@

Notice that it is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus didnot experiment.

@@

		$5 \times100$\  &   $16\%$ & $51\%$ & $\mathbf{69.9\%}$ & $57.5\%$ & $66.1\%$ & $65.3\%$ & $68.4\%$ &  $84.2 \%$ \\ 
		$\epsilon = 0.026$ & 5s & 159s & 102s& 75s & 140s & 117s & 142s &  \\
		\hline

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accuate than using compensation: we are within $1.6\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).
